Long-term cloud enhancement events of global solar irradiance over Thessaloniki, Greece

### 2024-06-22 article/article.Rmd
Commit 2024-06-22 19:11

---
title: Long term cloud enhancement events of global solar irradiance over Thessaloniki, Greece

@@

Instruments respose ....
@Martins2022
-->

@@

<!-- our work -->
In this study, we evaluate the effects CE on GHI by investigating their frequency of
occurrence, intensity, and duration in a thirty-year period of GHI observations at
Thessaloniki, Greece.  We used modeled clear sky irradiance as a baseline to identify
cloud enhancements and we determined long term trends of the above mentioned metrics,
their climatology and some general characteristics.  To our knowledge there are no
other studies that provide trends from such  long dataset.

@@

### Long term change of clear sky irradiance

<!-- Aerosol change -->
The above clear sky reference values are based on the climatological AOD and WC,
hence they cannot describe accurately the long term variation of
$\text{GHI}_\text{ref}$ due to changes in the two atmospheric constituents, mainly
AOD.  As reported by @Natsis2023, there is a long term brightening effect in the GHI
data of Thessaloniki for the period 1993 -- 2023, which for clear-sky data was
attributed to long-term changes in aerosol effects.  Therefore, an adjustment of the
$\text{GHI}_\text{ref}$ during the period of study was made using simulation with the
RTM based on the long-term variations of the AOD. As AERONET data start only in 2003,
we used for the period 1993 -- 2005 estimates of changes in AOD at $340\,\text{nm}$
derived from a collocated Brewer spectrophotometer [@Kazadzis2007] to calculate the
trend in $\text{GHI}_\text{ref}$ during this period.

@@

<!--
To create a unified trend for the long term change of the clear sky irradiance, we
simulated the values of AOD derived from those sources with Libradtran. For both
inputs, we used the AOD at $500\,\text{nm}$, which was inferred by the available
Ångström coefficients. We choose the SZA of $55^\circ$ as a representative value for
all the runs.
-->

According to this study, in the period 1997 -- 2005 the mean AOD at $340\,\text{nm}$
is $0.403$ with a trend of $-3.8\pm0.93\,\%$ per year, corresponding to a change of
$0.0153$ per year.  Using an Ångström coefficient $\alpha = 1.6$, this translates to
a change in the Ångström coefficient $\beta=0.00272$ per year (or $\beta=0.084$ in
1997 and $\beta=0.059$ in 2005).  Simulations with uvspec for the above Ångström
coefficients, with WC of
$`r round(BR[year==1997,pw_avg_mm],1)`\,\text{mm}$ and 
$`r round(BR[year==2005,pw_avg_mm],1)`\,\text{mm}$
for 1997 and 2005 respectively and for a SZA of $55^\circ$ reveal a trend of
$+0.21\,\%$ per year in $\text{GHI}_\text{ref}$. 
The SZA of $55^\circ$ was chosen as representative of all days in the year in order
to get a rough estimate of the annually averaged change in clear sky irradiance.  For
the period 2005 -- 2023 we used the mean monthly values of AOD and WC from AERONET in
a similar simulation scheme to calculate the monthly mean of clear sky irradiance,
and finally the trend of $+0.14\,\%$ per year. We applied these two long term changes
(see Figure\nobreakspace{}\ref{fig:CS-change}) to the climatological
$\text{GHI}_\text{ref}$, in order to create a more realistic representation of the
clear-sky irradiance for the whole period of study.

```{r CS-change, echo=FALSE, fig.cap="Simulated long term change in clear sky irradiance relative to the climatological values due to changes in AOD."}
knitr::include_graphics("../images/P-CS-change-1.pdf")
```

@@

In this study our main focus was to quantify over irradiance events related to CEs. A
key issue for achieving this goal is to  define a threshold for the CE
identification, representative of the clear-sky irradiance at the time of each GHI
measurement. This depends on the selection of the appropriate atmospheric
parameterization for the RTM simulations. The implementation of the long term change
of AOD, discussed in section\nobreakspace{}\ref{rtmcs}, allows capturing a large part
of the natural variability of clear-sky GHI. However, the short-term variability of
AOD cannot be taken adequately into account when using monthly values for the model
simulations. We tried different approaches to increase the robustness of the
methodology and compensate for the limited accuracy of the RTM input data and the
unpredictable natural variability of the atmosphere.

First, we evaluated the performance of the modelled $\text{GHI}_\text{ref}$ in
relation to the measured GHI for different levels of atmospheric clearness, by using
in the RTM the monthly climatological AOD and WC, less their respective standard
deviations.
<!-- $\tau_{\text{cs}} = \tau_{500\text{nm}} - 1\sigma$ and $w_{h\text{cs}} = w_h - 1\sigma$
($\text{GHI}_\text{CSm}$). -->
These values represent typical atmospheres in Thessaloniki with lower than average
load of aerosols and humidity, which are the main factors that attenuate the GHI,
excluding clouds.  With this approach the simulated $\text{GHI}_\text{ref}$ should be
generally greater than the measured GHI when aerosols are more abundant.  To
compensate for this, we increased the $\text{GHI}_\text{ref}$ by 
($`r 100*(C4_cs_ref_ratio-1)`\,\%$)
with an additional constant offset of
$`r C4_GLB_diff_THRES`\,\text{W}/\text{m}^2$,
as described in Equation\nobreakspace\ref{eq:CE4}.
\begin{equation}
\text{CE} : \text{E}_\text{i} > `r C4_GLB_diff_THRES` + `r C4_cs_ref_ratio` \cdot \text{E}_\text{CSm,i} \,\,[\text{W}/\text{m}^2] \label{eq:CE4}
\end{equation}
where: $\text{E}$ the measured irradiance, $\text{E}_\text{CSm}$ the selected
modelled clear sky irradiance, and $i$ each of the one-minute observation.  This is
the criterion of our CE identification.

These values were determined through the implementation of an empirical method with
manual inspection of the CE identification results  on selected days of the whole
dataset.  We used seven sets of days with characteristics relevant to the efficiency
of the identification threshold.  These sets were random groups of about 20 -- 30
days with the following characteristics:
(a) the strongest over irradiance CE events,
(b) the largest daily total over irradiance,
(c) absence of clouds (by implementing a clear sky identification algorithm as discussed in @Natsis2023),
(d) absence of clouds and absence of EC events,
(e) with at least $60\,\%$ of the day length without clouds and presence of EC events,
(h) randomly selected days, and
(i) manually selected days.
For the latter case and where needed in some of the edge cases, we also used images
from a sky-camera to further aid the decision of the manual inspection.

@@

The distribution of the number of CE events as a function of OIR is shown in
Figure\nobreakspace{}\ref{fig:ovir-distribution}. Apparently, there is an inverse
relation between the frequency of CE events  and OIR with an exponential-like
decline. This is expected, as the stronger the CE events become, the rarer are the
conditions favoring the occurrence of CE events. For the majority (over $62\,\%$) of
the CE events the OIR is below the long term average of
$`r round(mean(ST_E_yearly$GLB_diff.mean), 1)`\,\text{W}/\text{m}^2$,
while about $8.1\,\%$ of the events correspond to OIR larger than
$100\,\text{W}/\text{m}^2$ and up to highest value of $412.4\,\text{W}/\text{m}^2$.
This distribution is indicative of the magnitude and the probability of the expected
CE events over Thessaloniki.  Similar distribution of CE events, albeit with larger
OIR values, has been reported by @Vamvakas2020 for the city of Patras. This site is
located $2.5^\circ$ south of Thessaloniki and is exposed to air masses coming mainly
from the eastern Mediterranean resulting in different cloud patterns, that may affect
the characteristics and magnitude of the CE events.

>>>>

---
title: Long-term cloud enhancement events of global solar irradiance over Thessaloniki, Greece

@@

Instruments response ....
@Martins2022
-->

@@

<!-- our work -->
In this study, we evaluate the effects CE on GHI by investigating their frequency of
occurrence, intensity, and duration in a thirty-year period of GHI observations at
Thessaloniki, Greece.  We used modeled clear sky irradiance as a baseline to identify
cloud enhancements and we determined long-term trends of the above mentioned metrics,
their climatology and some general characteristics.  To our knowledge there are no
other studies that provide trends from such  long dataset.

@@

### Long-term change of clear sky irradiance

<!-- Aerosol change -->
The above clear sky reference values are based on the climatological AOD and WC,
hence they cannot describe accurately the long-term variation of
$\text{GHI}_\text{ref}$ due to changes in the two atmospheric constituents, mainly
AOD.  As reported by @Natsis2023, there is a long-term brightening effect in the GHI
data of Thessaloniki for the period 1993 -- 2023, which for clear-sky data was
attributed to long-term changes in aerosol effects.  Therefore, an adjustment of the
$\text{GHI}_\text{ref}$ during the period of study was made using simulation with the
RTM based on the long-term variations of the AOD. As AERONET data start only in 2003,
we used for the period 1993 -- 2005 estimates of changes in AOD at $340\,\text{nm}$
derived from a collocated Brewer spectrophotometer [@Kazadzis2007] to calculate the
trend in $\text{GHI}_\text{ref}$ during this period.

@@

<!--
To create a unified trend for the long-term change of the clear sky irradiance, we
simulated the values of AOD derived from those sources with Libradtran. For both
inputs, we used the AOD at $500\,\text{nm}$, which was inferred by the available
Ångström coefficients. We choose the SZA of $55^\circ$ as a representative value for
all the runs.
-->

According to this study, in the period 1997 -- 2005 the mean AOD at $340\,\text{nm}$
is $0.403$ with a trend of $-3.8\pm0.93\,\%$ per year, corresponding to a change of
$0.0153$ per year.  Using an Ångström coefficient $\alpha = 1.6$, this translates to
a change in the Ångström coefficient $\beta=0.00272$ per year (or $\beta=0.084$ in
1997 and $\beta=0.059$ in 2005).  Simulations with uvspec for the above Ångström
coefficients, with WC of
$`r round(BR[year==1997,pw_avg_mm],1)`\,\text{mm}$ and 
$`r round(BR[year==2005,pw_avg_mm],1)`\,\text{mm}$
for 1997 and 2005 respectively, also taken from AERONET, and for a SZA of $55^\circ$
reveal a trend of
$+0.21\,\%$ per year in $\text{GHI}_\text{ref}$. 
The SZA of $55^\circ$ was chosen as representative of all days in the year in order
to get a rough estimate of the annually averaged change in clear sky irradiance.  For
the period 2005 -- 2023 we used the mean monthly values of AOD and WC from AERONET in
a similar simulation scheme to calculate the monthly mean of clear sky irradiance,
and finally the trend of $+0.14\,\%$ per year. We applied these two long-term changes
(see Figure\nobreakspace{}\ref{fig:CS-change}) to the climatological
$\text{GHI}_\text{ref}$, in order to create a more realistic representation of the
clear-sky irradiance for the whole period of study.

```{r CS-change, echo=FALSE, fig.cap="Simulated long-term change in clear sky irradiance relative to the climatological values due to changes in AOD."}
knitr::include_graphics("../images/P-CS-change-1.pdf")
```

@@

In this study our main focus was to quantify over irradiance events related to CEs. A
key issue for achieving this goal is to  define a threshold for the CE
identification, representative of the clear-sky irradiance at the time of each GHI
measurement. This depends on the selection of the appropriate atmospheric
parameterization for the RTM simulations. The implementation of the long-term change
of AOD, discussed in section\nobreakspace{}\ref{rtmcs}, allows capturing a large part
of the natural variability of clear-sky GHI. However, the short-term variability of
AOD cannot be taken adequately into account when using monthly values in the model
simulations. We tried different approaches in order to strengthen the robustness of
the methodology and compensate for the limited accuracy of the RTM input data and the
unpredictable natural variability of the atmosphere.

First, we evaluated the performance of the modelled $\text{GHI}_\text{ref}$ in
relation to the measured GHI for different levels of atmospheric clearness, by using
in the RTM the monthly climatological AOD and WC, less their respective standard
deviations.
<!--
$\tau_{\text{cs}} = \tau_{500\text{nm}} - 1\sigma$ and $w_{h\text{cs}} = w_h - 1\sigma$
($\text{GHI}_\text{CSm}$).
-->
These values represent typical atmospheres in Thessaloniki with lower than average
load of aerosols and humidity, which are the main factors that attenuate the GHI,
excluding clouds.  With this approach the simulated $\text{GHI}_\text{ref}$ should be
generally greater than the measured GHI when aerosols are more abundant.  To
compensate for this, we defined the following threshold $\text{E}$ to compare the
measured $\text{GHI}$: 
<!--
($`r 100*(C4_cs_ref_ratio-1)`\,\%$)
with an additional constant offset of
$`r C4_GLB_diff_THRES`\,\text{W}/\text{m}^2$,
as described in Equation\nobreakspace\ref{eq:CE4}.
-->
\begin{equation}
\text{CE} : \text{E} > `r C4_GLB_diff_THRES` + `r C4_cs_ref_ratio` \cdot \text{GHI}_\text{ref} \,\,[\text{W}/\text{m}^2] \label{eq:CE4}
\end{equation}
<!--
where: $\text{E}$ the measured irradiance, $\text{E}_\text{CSm}$ the selected
modelled clear sky irradiance, and $i$ each of the one-minute observation.  This is
the criterion of our CE identification.
-->
This is the criterion of our CE identification.

These values were determined through the implementation of an empirical method with
manual inspection of the CE identification results  on selected days of the whole
dataset.  We used seven sets of days with characteristics relevant to the efficiency
of the identification threshold.  These sets were random groups of about 20 -- 30
days with the following characteristics:
(a) the strongest over irradiance CE events,
(b) the largest daily total over irradiance,
(c) absence of clouds (by implementing a clear sky identification algorithm as discussed in @Natsis2023),
(d) absence of clouds and absence of EC events,
(e) with at least $60\,\%$ of the day length without clouds and presence of EC events,
(h) randomly selected days, and
(i) manually selected days.
For the latter case and where needed, we also used images from a sky-camera to
further aid the decision of the manual inspection.

@@

The distribution of the number of CE events as a function of OIR is shown in
Figure\nobreakspace{}\ref{fig:ovir-distribution}. Apparently, there is an inverse
relation between the frequency of CE events  and OIR with an exponential-like
decline. This is expected, as the stronger the CE events become, the rarer are the
conditions favoring the occurrence of CE events. For the majority (over $62\,\%$) of
the CE events the OIR is below the longnterm average of
$`r round(mean(ST_E_yearly$GLB_diff.mean), 1)`\,\text{W}/\text{m}^2$,
while about $8.1\,\%$ of the events correspond to OIR larger than
$100\,\text{W}/\text{m}^2$ and up to highest value of $412.4\,\text{W}/\text{m}^2$.
This distribution is indicative of the magnitude and the probability of the expected
CE events over Thessaloniki.  Similar distribution of CE events, albeit with larger
OIR values, has been reported by @Vamvakas2020 for the city of Patras. This site is
located $2.5^\circ$ south of Thessaloniki and is exposed to air masses coming mainly
from the eastern Mediterranean resulting in different cloud patterns, that may affect
the characteristics and magnitude of the CE events.
